---
editor_options:
  chunk_output_type: console
---

# Speed and collar temperature

Here we look at the relationship between elephant speed and collar temperature, as well as some other relevant factors that we think might influence elephant movement.

## Load libraries

```{r load_libs_04, message=FALSE, warning=FALSE, eval=FALSE}

# load libraries
library(dplyr)
library(purrr)
library(lubridate)
library(readr)

# stats
library(mgcv)

# spatial
library(sf)
library(raster)

# plotting
library(ggplot2)
library(ggthemes)
library(viridis)

# custom funcs
ci = function(x) 1.96*sd(x, na.rm = T)/sqrt(length(x))
```

## Load data from sources

Load the tracking data and the transformed slope data.

```{r read_eles_04, eval=FALSE}
# read in eles as csv
data = read_csv("data/elephant_data.csv")
```

## Speed and temperature

### Run GAMM speed ~ temperature

```{r model_speed_temp, eval=FALSE}
# make id a factor
data$id = as.factor(data$id)
data$season = as.factor(data$season)

# run a GAMM using the mgcv package
mod.speed = bam(v ~ s(temp, k = 4) + 
  season + 
  woody.density + 
  s(id, bs = "re") + 
  s(hour, bs = "re"), 
  data = data)

# print model summary to file
if(!dir.exists("data/model_output")){
  dir.create("data/model_output")
}

if(file.exists("data/model_output/model_gamm_speed.txt")){
  file.remove("data/model_output/model_gamm_speed.txt")
}

R.utils::captureOutput(summary(mod.speed), 
              file = "data/model_output/model_gamm_speed.txt",
              append = TRUE)
```

Print the GAMM summary.

```{r show_model_speed, eval=TRUE, comment=''}
cat(readLines("data/model_output/model_gamm_speed.txt"), sep = "\n")
```

### Prepare speed ~ temp plot data

```{r ele.speed.temp.data, eval=FALSE}
# prepare data for plotting
ele.speed.temp = 
  data %>%
  mutate(v.pred = predict(mod.speed, 
                          newdata = ., scale = "response", 
                          allow.new.levels = T), 
         temp = plyr::round_any(temp,2)) %>%
  ungroup() %>% 
  
  group_by(season, temp) %>%
  
  summarise(v.mean = mean(v), 
            v.sd = sd(v), 
            n.v = length(v), 
            pred.mean = mean(v.pred, na.rm = T), 
            pred.sd=sd(v.pred, na.rm = T), 
            pred.n = length(v.pred)) %>%
  
  mutate(v.ci = qnorm(0.975)*v.sd/sqrt(n.v), 
         ci.pred = qnorm(0.975)*pred.sd/sqrt(pred.n))
```

### Prepare Figure 4 (A): Speed and temperature

```{r fig_speed_temp, eval=FALSE}
# figure for speed and temperature
fig_speed_temp =
  ele.speed.temp %>% 
  filter(temp %in% 10:40) %>%
  ggplot()+
    
  geom_rangeframe(data = data_frame(x=c(10,40), y = c(0.15,0.6)),aes(x,y))+
  geom_smooth(aes(x = temp, y = pred.mean*2/1e3, 
                  col = season, fill = season, lty = season), 
              alpha = 0.2, lwd = 0.5)+
  geom_pointrange(aes(x = temp, y = v.mean*2/1e3, 
                      ymin = (v.mean-v.ci)*2/1e3, ymax = (v.mean+v.ci)*2/1e3, 
                      col = season, shape = season), 
                  fill = "white", size = 0.4, stroke =0.7, lty = 1, 
                  position = position_dodge(width = 0.3))+

  scale_fill_brewer(palette = "Set1")+
  scale_color_brewer(palette = "Set1")+
  scale_shape_manual(values=c(21,24))+
  scale_linetype_manual(values=c("dashed","solid"))+
  
  theme_few()+
  theme(panel.border = element_blank(),
        legend.position = "none")+
  coord_cartesian(ylim=c(0.15,0.6))+
  scale_y_continuous(breaks = seq(.15,.6,.15))+
  labs(x = "Collar temperature (°C)", y = "Speed (km/h)",
            col = "Season", fill = "Season", title="(a)")

```

### Prepare data for speed ~ woodland plot

```{r speed_wood_plots, eval=FALSE}
# prepare data
ele.speed.wood = 
  as_data_frame(data) %>% 
  mutate(v.pred = predict(mod.speed, 
    newdata = ., scale = "response", allow.new.levels = T)) %>%  
  
  dplyr::select(woody.density, v, v.pred, season) %>% 
  mutate(v2 = v*2/1e3, 
    v.pred2 = v.pred*2/1e3) %>% 
  
  dplyr::select(-v, -v.pred) %>% 
  
  tidyr::gather(var, value, -woody.density, -season) %>% 
  group_by(season, wood = plyr::round_any(woody.density, 5),var) %>%
  
  summarise_at(vars(value), list(mean=mean, sd=sd, length=length)) %>% 
  mutate(ci = 1.96*sd/sqrt(length))

```

### Prepare Figure 4 (B): Speed and woody cover

```{r fig_speed_wood, eval=FALSE}
#review figs: speed vs slope, speed vs woody density
fig_speed_wood = ggplot()+
  geom_smooth(data = ele.speed.wood %>% filter(var == "v.pred2"), 
              aes(x = wood, y = mean, col = season, fill = season, lty = season), 
              alpha = 0.2, size = 0.3)+
  
  geom_pointrange(data = ele.speed.wood %>% filter(var == "v2"), 
                  aes(x = wood, ymin = mean-ci, ymax = mean+ci, y = mean, 
                      col = season, shape = season), 
                  fill = "white", position = position_dodge(width = 0.3), 
                  size = 0.4, stroke = 0.7)+
  
  geom_rangeframe(data = data_frame(x=c(0,80),y=c(0.15,0.6)), aes(x,y))+
  
  #facet_wrap(~var_name, scales = "free_x")+
  
  scale_color_brewer(palette = "Set1")+
  scale_fill_brewer(palette = "Set1")+
  scale_linetype_manual(values=c(2,1))+
  scale_shape_manual(values= c(21,24))+
  
  theme_few()+
  theme(panel.border = element_blank(),
        legend.position = "none")+
  labs(x = "Woody cover (%)", y = "Speed (km/h)", title = "(b)")+
  
  scale_x_continuous(breaks = seq(0, 80, 20))+
  scale_y_continuous(breaks = c(.15,.3,.45,.6), limits = c(NA, .6))+
  coord_cartesian(ylim=c(.15,.6))
```

## Figure 4: Speed and temperature and woody cover

```{r prep_figure_speed_predictors, eval=FALSE}
half = 85/25.4; full = 180/25.4

# export fig for temp measures
library(gridExtra)

figure_04_speed = grid.arrange(fig_speed_temp, fig_speed_wood, nrow = 1)

ggsave(figure_04_speed,
       filename = "figs/figure_04_speed.pdf",
       height = half, width = full)
```

```{r show_figure_04, eval=TRUE, echo=TRUE, message=FALSE, fig.cap="Speed of elephant movement in relation to **(a)** collar temperature (at 2 ◦ C intervals) and **(b)** % woody cover (at 5 unit intervals) in the dry (red circles) and wet season (blue triangles). GAMM fit (lines) and 95% confidence intervals (vertical line ranges and shaded areas) are shown for each season separately."}
library(magick)
fig04 = magick::image_read_pdf("figs/figure_04_speed.pdf")
fig04
```