
# identifying loops

```{r}
source("libs.R")
source("ggplot.opts.r")
source("ggplot.pub.r")
```

```{r}
load("ele.revisits.rdata")
load("eledata.rdata")

ele.rev = left_join(ele.rev, ele2)

```

# find high rev and low distance to water

```{r}
#'quantiles
quantile(ele.rev$angle, na.rm = T)

ele.rev$watervisits = ifelse(ele.rev$mindw <=500,# & ele.rev$residence >=10 & ele.rev$revisits >= 10, 
1, 0)

#ggplot(ele.rev)+ geom_smooth(aes(x = mindw, y = residence/revisits))+
  #coord_cartesian(xlim = c(0,5e2), ylim = c(1.5,1.65))

#a = ele.rev %>% filter(watervisits == 1, !is.na(waterdiff)) %>% select(id, xutm, yutm, v, angle, mindw, revisits, residence, fpt, waterdiff)

#write.csv(a, file = "ele.watervisits.csv", row.names = F)
```

# find water visits by each ele

```{r}
#'separate by ele, and find the time between consecutive watervisits

ele3 = ele.rev %>% dlply("id") %>%
  map(function(x) x %>% filter(watervisits == 1))

#'here, use diff rather than difftime, adding NA for the first point, and getting the output in numeric hours
for(i in 1:14){
  ele3[[i]]$wvint = c(NA, as.numeric(diff(ele3[[i]]$time)/60))
}

ele3 = bind_rows(ele3)

```

# plot time between water visits

```{r}
#'remove bad difftimes.
#### don't filter just yet####
ele4 = ele3 #%>% filter(wvint > 3, wvint < 72)
```

## Prep interval distribution

```{r interval_distr}

#'plot to see dist of revisit rates
#fig2b = 
ggplot()+
  stat_density(data = ele3 %>% filter(wvint > 3, wvint < 72),aes(x = wvint, col = season2), position = "identity",geom = "line", lwd = 0.7)+
  geom_rect(aes(xmin = 12, xmax = 24, ymin = 0.0, ymax = 0.035), col = "grey20", lwd = 0.2, fill = "transparent")+
 scale_colour_manual(values = c(cola, colb))+
  #facet_wrap(~ifelse(season2 == "dry", "Cool dry", "Hot wet"))+
  theme_pub()+
  labs(list(x = "Water visit interval (hrs)", title = "(b)"))
```

# Distance between track start and end

```{r}
#'remove points where the water visit interval is less than 3, these are points where the elephant is loitering near water. these are water visit points

#### dont' filter just yet ####

#ele4 = ele3 %>% filter(wvint > 3, wvint < 72)

#'select all points which are not water visits
ele.rev2 = anti_join(ele.rev, ele4, by = c("id", "time"))

#'assign a watervisit identifier of 0, or false, and a watervisit interval of NA
ele.rev2$watervisits = 0; ele.rev2$wvint = NA

#'recombine the non-visit and visit points, it should be the same size as the base data, here, 278757
ele5 = rbind(ele.rev2, ele4)

#'sort the recombined data by id and time
ele5 = ele5 %>% arrange(id, time)
```

## identifying and numbering segments between water

```{r}
#'split by elephant and identify the segments between water
ele5  = ele5 %>% dlply("id") %>%
  map(function(x){
    x %>% mutate(loop = cumsum(watervisits))
  })

#'for each elephant and each segment, find
#'1. the time difference between the ith point of the segment and the start
#'2. the change in distance to water from the start point
#'3. the change in temperature since the start
#'4. the instantaneous change in temperature
#'5. the proportion of the segment completed at each point
ele6 =  ele5 %>%
  map(function(x){
    x %>% group_by(loop) %>% mutate(looptime = as.numeric(difftime(time, time[1], units = "hours")), deltadw = mindw - mindw[1], tempchange = temp - temp[1], ambtempdiff = c(NA, diff(temp))) %>% #filter(looptime < 24) %>% 
      mutate(loopprop = looptime/max(looptime))
  })

#'now find the displacement between the first and last point
library(sp)
ele6 = ele6 %>% bind_rows()

#'get the first and last x-y coords, the season, the time taken for the loop to be completed, the mean speed, the temperature at the halfway point, the mean temperature along the loop, the instantaneous change in temp at the halfway stage
ele7 = ele6 %>% group_by(id, loop) %>% arrange(id, loop, time) %>% 
  summarise(x1 = first(xutm), 
            y1 = first(yutm), 
            x2 = last(xutm), 
            y2 = last(yutm),
            long2 = last(long),
            lat2 = last(lat),
            season = first(season2), 
            looptime = max(looptime, na.rm = T), 
            v = mean(v, na.rm = T), 
            t50 = temp[first(which(round_any(loopprop, 0.05) == 0.5))], 
            temp = mean(temp, na.rm=T), 
            tdiff50 = ambtempdiff[first(which(round_any(loopprop, 0.05) == 0.5))], 
            wvint = first(wvint), 
            distance = sum(distance, na.rm=T), 
            maxdw = max(mindw,na.rm = T), 
            hour_start = first(hour), 
            hour_50 = hour[first(which(round_any(loopprop, 0.05) == 0.5))], hour_end = last(hour), 
            mdw_start = first(mindw), 
            mdw_end = last(mindw), 
            x50 = long[first(which(round_any(loopprop, 0.05) == 0.5))], 
            y50 = lat[first(which(round_any(loopprop, 0.05) == 0.5))],
            time50 = time[first(which(round_any(loopprop, 0.05) == 0.5))])

ele.all.loops = ele7

#'remove loops with a looptime above 72
ele7 = ele7 %>% filter(looptime <=72, looptime > 3)

#'split by each loop
ele7 = ele7 %>% dlply(c("id","loop"))

#'now find distances between first and last points of the loop
for(i in 1:length(ele7)){
  a = ele7[[i]]
  ele7[[i]]$between_point_dist = (spDists(x = as.matrix(a[c("x1","y1")]), y = as.matrix(a[c("x2","y2")])))
}

ele7 = ele7 %>% rbind_list()
```

## histogram of durations

**this is the only plot looking at all segments**

```{r}
#pdf(file = "duration.density.pdf", width = 5,height = 3)
fig5a = ggplot()+
  stat_density(data = ele7, aes(x = looptime, y = ..density.., lty = season),lwd = 0.5, geom = "line", position = "identity")+
  geom_rect(aes(xmin = 12, xmax = 24, ymin = -0.001, ymax = 0.042), col = "grey20", lwd = 0.2, fill = "transparent")+
 #scale_colour_manual(values = c(colb3, colb3))+
  scale_linetype_manual(values = c(2,1))+
# scale_colour_manual(values = c(1,"transparent"))+
  #facet_wrap(~ifelse(season2 == "dry", "Cool dry", "Hot wet"))+
  theme_pub()+
  labs(list(x = "Segment duration (hours)", y = "Probability density", title = "(a)"))

fig5a.col =
  ggplot()+
  stat_density(data = ele7, aes(x = looptime, y = ..density.., lty = season, col = season),lwd = 0.5, geom = "line", position = "identity")+
  geom_rect(aes(xmin = 12, xmax = 24, ymin = -0.001, ymax = 0.042), col = "grey20", lwd = 0.2, fill = "transparent")+
 #scale_colour_manual(values = c(colb3, colb3))+
  scale_linetype_manual(values = c(2,1))+
  scale_colour_manual(values = c(cola2, colb2))+
  #facet_wrap(~ifelse(season2 == "dry", "Cool dry", "Hot wet"))+
  theme_pub()+
  labs(list(x = "Segment duration (hours)", y = "Probability density", title = "(a)"))

#dev.off()
```

## how far are midpoints from each other?

```{r}
#'find the distance between 1st midpoint and all other midpoints
#'first filter nas
ele8 = ele7 %>% filter(!is.na(x50), !is.na(y50), looptime>12, looptime< 24) %>% arrange(id, time50) %>% 
  dlply("id")

library(geosphere)
ele_p50_dists = ele8 %>% 
  map(function(x){data_frame(
   p50dist =  c(NA, distVincentyEllipsoid(p1 = x[c("x50","y50")]#, 
                                 # p2 = x[-1,c("x50","y50")]
                                )),
   p_end_dist = c(NA, distVincentyEllipsoid(p1 = x[c("long2","lat2")])),
   loop = x$loop,
   id = x$id,
   season = x$season)
  }) %>% bind_rows()

```

**ALL FURTHER PLOTS SHOW ONLY 12 - 24 HOUR SEGMENTS**

  ## plots of midpoint distance
  
```{r}
  #pdf(file = "midpoint.distance.density.pdf", width = 5, height = 3)
  
  #ele_p50_dists %>% 
   # ggplot()+
   # geom_point(aes(x = p_end_dist, y = p50dist, col = season))+
   # xlim(0,10e3)+ylim(0,10e3)
  
  fig5d = ele_p50_dists %>% 
    ggplot()+
    stat_density(aes(x = p50dist/1e3, y = ..density.., lty = season), position = "identity", lwd = 0.5, geom = "line")+
 # scale_colour_manual(values = c(cola3, colb3))+
 #scale_colour_manual(values = c(1,"transparent"))+
  scale_linetype_manual(values = c(2,1))+
  theme_pub()+
  xlim(0,10)+
  labs(list(x = "Displacement segment midpoints (km)", y = "Probability density", title = "(d)"))

#'ele midpoint distances under 1000m
#count(ele_p50_dists, p50dist < 1000)

#'ele displacements under 1000m
#count(ele7, between_point_dist < 1000)
#dev.off()

 fig5d.col = ele_p50_dists %>% 
    ggplot()+
    stat_density(aes(x = p50dist/1e3, y = ..density.., lty = season, col = season), position = "identity", lwd = 0.5, geom = "line")+
 scale_colour_manual(values = c(cola2, colb2))+
 #scale_colour_manual(values = c(1,"transparent"))+
  scale_linetype_manual(values = c(2,1))+
  theme_pub()+
  xlim(0,10)+
  labs(list(x = "Displacement segment midpoints (km)", y = "Probability density", title = "(d)"))


```



## distance along segment

```{r}
#pdf(file = "segment.length.density.pdf", width = 5, height = 3)

fig5b = ggplot()+
  stat_density(data = ele7 %>% filter(looptime >12, looptime < 24), aes(x = distance/1e3, y = ..density.., lty = season), position = "identity",lwd = 0.5, geom = "line")+
  #scale_colour_manual(values = c(cola3, colb3))+
  #scale_fill_manual(values = c(1, "transparent"))+
  #facet_wrap(~ifelse(season2 == "dry", "Cool dry", "Hot wet"))+
  scale_linetype_manual(values = c(2,1))+
  theme_pub()+
  labs(list(x = "Segment length (km)", y = "Probability density", title = "(b)"))

#dev.off()

fig5b.col = ggplot()+
  stat_density(data = ele7 %>% filter(looptime >12, looptime < 24), aes(x = distance/1e3, y = ..density.., lty = season, col = season), position = "identity",lwd = 0.5, geom = "line")+
  scale_colour_manual(values = c(cola2, colb2))+
  #scale_fill_manual(values = c(1, "transparent"))+
  #facet_wrap(~ifelse(season2 == "dry", "Cool dry", "Hot wet"))+
  scale_linetype_manual(values = c(2,1))+
  theme_pub()+
  labs(list(x = "Segment length (km)", y = "Probability density", title = "(b)"))
```


## displacement density plot

```{r ele_looping}
#pdf(file = "displacement.density.pdf", height = 3, width = 5)
#fig2a = 
fig5c = ggplot(ele7 %>% filter(looptime > 12, looptime < 24))+
  stat_density(aes(x = between_point_dist/1e3, y = ..density.., lty = season), position = "identity", geom = "line")+
 # scale_colour_manual(values = c(cola3, colb3))+
  scale_linetype_manual(values = c(2,1))+
  theme_pub()+
  labs(list(x = "Displacement start & end (km)", y = "Probability density", title = "(c)"))+
  xlim(0,10)

#dev.off()

fig5c.col = ggplot(ele7 %>% filter(looptime > 12, looptime < 24))+
  stat_density(aes(x = between_point_dist/1e3, y = ..density.., lty = season, col = season), position = "identity", geom = "line")+
  scale_colour_manual(values = c(cola2, colb2))+
  scale_linetype_manual(values = c(2,1))+
  theme_pub()+
  labs(list(x = "Displacement start & end (km)", y = "Probability density", title = "(c)"))+
  xlim(0,10)

```

## export panel figure 5

```{r}
pdf(file = "segment.dist.disp.dens.pdf", height = 6, width = 6)

library(gridExtra)

grid.arrange(fig5a,fig5b, fig5c, fig5d, ncol = 2)

dev.off()

```

## export fig 5 colour

```{r}
pdf(file = "segment.dist.disp.dens.colour.pdf", height = 6, width = 6)

library(gridExtra)

grid.arrange(fig5a.col,fig5b.col, fig5c.col, fig5d.col, ncol = 2)

dev.off()
```


## distance and displacement

## b/w plot
```{r}
#'export
pdf(file = "displacement.distance.pdf", width = 5, height = 3)

ele7 %>% filter(looptime> 12, looptime < 24) %>% 
  group_by(mdist = round_any(distance, 500), season) %>% 
  summarise(mdisp = mean(between_point_dist, na.rm = T), sd = sd(between_point_dist, na.rm = T), n = length(between_point_dist)) %>% 
  mutate(ci = qnorm(0.975)*sd/sqrt(n)) %>% 
  
  ggplot(aes(x = mdist/1e3, y = mdisp/1e3, group = season))+
  
  geom_smooth(data = ele7 %>% filter(looptime > 12, looptime < 24),  aes(x = distance/1e3, y = between_point_dist/1e3, lty = season), col = 1,  method = "glm", formula = y~x, family = "gaussian", lwd = 0.5)+
  
  geom_pointrange(aes(ymin = (mdisp-ci)/1e3,ymax = (mdisp+ci)/1e3, shape = season), fill = "white", position = position_dodge(width = 0.8), stroke = 0.6)+
# facet_wrap(~dur, scales = "free")+
  theme_pub()+theme(strip.text = element_text(colour = 1))+
  scale_shape_manual(values = c(21,16))+
  scale_linetype_manual(values = c(2,1))+
  labs(list(x = "Distance along segment (km)", y = "Displacement segment start & end (km)"))+
  #xlim(1.25,10)+ ylim(0,5)
  coord_cartesian(xlim = c(1.5,10), ylim = c(0,5))

dev.off()
```

### colour plot

```{r}
pdf(file = "displacement.distance.colour.pdf", width = 5, height = 3)

ele7 %>% filter(looptime> 12, looptime < 24) %>% 
  group_by(mdist = round_any(distance, 500), season) %>% 
  summarise(mdisp = mean(between_point_dist, na.rm = T), sd = sd(between_point_dist, na.rm = T), n = length(between_point_dist)) %>% 
  mutate(ci = qnorm(0.975)*sd/sqrt(n)) %>% 
  
  ggplot(aes(x = mdist/1e3, y = mdisp/1e3, group = season))+
  
  geom_smooth(data = ele7 %>% filter(looptime > 12, looptime < 24),  aes(x = distance/1e3, y = between_point_dist/1e3, lty = season, col = season, fill = season),  method = "glm", formula = y~x, family = "gaussian", lwd = 0.5)+
  
  geom_pointrange(aes(ymin = (mdisp-ci)/1e3,ymax = (mdisp+ci)/1e3, shape = season, col = season), fill = "white", position = position_dodge(width = 0.8), stroke = 0.6)+
# facet_wrap(~dur, scales = "free")+
  theme_pub()+theme(strip.text = element_text(colour = 1))+
  scale_shape_manual(values = c(21,16))+
  scale_linetype_manual(values = c(2,1))+
  scale_colour_manual(values = c(cola2,colb2))+
  scale_fill_manual(values = c(cola1,colb1))+
  labs(list(x = "Distance along segment (km)", y = "Displacement segment start & end (km)"))+
  #xlim(1.25,10)+ ylim(0,5)
  coord_cartesian(xlim = c(1.5,10), ylim = c(0,5))

dev.off()
```


```{r}
#'get the day loops
dayloops = ele7 %>% filter(looptime > 12, looptime<24)

#'now get only the loop and id data associated with these loops from ele6, which is all points

dayloopdata = ele7 %>% filter(looptime > 12, looptime<24) %>% select(id, loop) %>% left_join(ele6, by = c("id", "loop"))
```


## whats's happening in loops?

```{r}
#'loop data summary
dayloopdata = dayloopdata %>%
  #bind_rows() %>%
  #filter(looptime < 24, looptime > 12) %>%
  select(loopprop, mindw, waterdiff, temp, fpt, loop, season2, v, tempchange, woody.density, hour, ambtempdiff)
```

# plots

### stage and time of day

```{r}
#'use viridis throughout? no, red blue is understood by more people
#'purple is cool dry,green is hot wet
library(viridis)
#fig loop1 time ~ prop
#figS4 =
  dayloopdata %>%
    group_by(season2, loopprop = as.factor(round_any(loopprop, 0.1))) %>%
    #summarise(meanhour = mean(hour)) %>%
  ggplot()+
    theme_pub()+

#  geom_smooth(aes(x = loopprop, y = hour, col = season2, fill = season2), lty = 1, se = T, lwd = 0.5)+
 # geom_violin(aes(x = loopprop, y = hour), fill = "grey90", col = "transparent", position = position_dodge(width = 0.95))+
    geom_boxplot(aes(x = loopprop, y = hour, fill = season2, col = season2), position = position_dodge(width = 0.95), lwd = 0.3, notch = F, outlier.size = 0.2, outlier.colour = 1, outlier.shape = 21)+
    
  scale_color_manual(values = c(cola, colb))+
  scale_fill_manual(values = c(cola2, colb2))+
  scale_x_discrete(breaks = seq(0,1,0.5), labels = c("Start", "50%", "End"))+
    labs(list(x = "Track stage", y = "Hour of day"))
```

## start, end, 50% plots

```{r}
seg_time = ele7 %>%  filter(looptime > 12, looptime<24) %>% 
  group_by(loop, season) %>% 
  melt(measure.vars =c("hour_50","hour_end","hour_start"))

seg_time$variable = revalue(seg_time$variable, replace = c("hour_50" = "Halfway", "hour_end" = "End", "hour_start" = "Start"))
seg_time$variable = factor(seg_time$variable, levels(seg_time$variable)[c(3,1,2)])
```

## segment stage and hour of day violin plot

```{r}
source("geom_flat_violin.r")
pdf(file="fig.seg_time.pdf", width = 5, height = 3)
ggplot()+
  geom_flat_violin(data = seg_time, aes(x = variable, y = value), col = "transparent", position = "identity",fill = "grey60", lwd = 0.5, draw_quantiles = c(0.5), scale = "width")+
  geom_boxplot(data = seg_time, aes(x = variable, y = value), width = 0.1)+
 # scale_fill_manual(values = c(cola3, colb3, 1))+
  labs(list(x = "Segment stage", y = "Hour of day"))+
  theme_pub()+
  coord_flip()

dev.off()

```


## prep data for ci plots

```{r}
dayloopsum = dayloopdata %>% 
  group_by(season2, loopprop = round_any(loopprop, 0.1)) %>% 
  select(loopprop, season2, mindw, temp, v, woody.density) %>% 
  melt(id.vars = c("loopprop","season2")) %>% 
  group_by(loopprop, season2, variable) %>% 
  summarise(mean = mean(value, na.rm = T), sd = sd(value, na.rm = T), n = length(value)) %>% 
  mutate(ci = qnorm(0.975)*sd/sqrt(n))
```

### plot all vars at once

```{r}
#'plot all
dayloopsum %>% filter(variable != "woody.density") %>% 

ggplot()+
  geom_pointrange(aes(x = loopprop, y = mean, ymin = mean-ci, ymax = mean+ci,  col = season2, shape = season2, fill = season2), lty = 1, lwd = 0.5, fatten = 5, stroke = 0.6, position = position_dodge(width = 0.1))+
  scale_color_manual(values = c(cola, colb))+
   scale_fill_manual(values = c(cola2, colb2))+
  scale_shape_manual(values = c(21,24))+
    labs(list(x = "Track stage", y = "Value"))+
  scale_x_continuous(breaks = seq(0,1,0.5), labels = c("Start", paste(50,"%", sep=""), "End"))+
  #facet_wrap(~variable, scales = "free")+
  theme_pub()
```

### Fig 2(c) Distance to water vs track stage

```{r mindw_loop}

fig_seg2 = 
  dayloopsum %>% filter(variable ==  "temp") %>% 

ggplot()+
  geom_pointrange(aes(x = loopprop, y = mean, ymin = mean-ci, ymax = mean+ci, shape = season2),  lty = 1, lwd = 0.5, fatten = 5, stroke = 0.4, position = position_dodge(width = 0.1), fill = "white")+
#  scale_color_manual(values = c(cola, colb))+
#   scale_fill_manual(values = c(cola2, colb2))+
  scale_shape_manual(values = c(21,16))+
    labs(list(x = "Track stage", y = "Temperature (째C)", title = "(b)"))+
  scale_x_continuous(breaks = seq(0,1,0.5), labels = c("Start", paste(50,"%", sep=""), "End"))+
  #facet_wrap(~variable, scales = "free")+
  theme_pub()

###

fig_seg2.col = 
  dayloopsum %>% filter(variable ==  "temp") %>% 

ggplot()+
  geom_pointrange(aes(x = loopprop, y = mean, ymin = mean-ci, ymax = mean+ci, shape = season2, col = season2),  lty = 1, lwd = 0.5, fatten = 5, stroke = 0.4, position = position_dodge(width = 0.1), fill = "white")+
    scale_color_manual(values = c(cola2, colb2))+
#   scale_fill_manual(values = c(cola2, colb2))+
  scale_shape_manual(values = c(21,16))+
    labs(list(x = "Track stage", y = "Temperature (째C)", title = "(b)"))+
  scale_x_continuous(breaks = seq(0,1,0.5), labels = c("Start", paste(50,"%", sep=""), "End"))+
  #facet_wrap(~variable, scales = "free")+
  theme_pub()


```

### Fig 2(e) Speed vs track stage

```{r}
fig_seg3 = 
  dayloopsum %>% filter(variable ==  "v") %>% 

ggplot()+
  geom_pointrange(aes(x = loopprop, y = mean*2, ymin = (mean-ci)*2, ymax = (mean+ci)*2, shape = season2), fill = "white", lty = 1, lwd = 0.5, fatten = 5, stroke = 0.4, position = position_dodge(width = 0.1))+
#  scale_color_manual(values = c(cola, colb))+
 #  scale_fill_manual(values = c(cola2, colb2))+
  scale_shape_manual(values = c(21,16))+
    labs(list(x = "Track stage", y = "Speed (m/hr)", title = "(c)"))+
  scale_x_continuous(breaks = seq(0,1,0.5), labels = c("Start", paste(50,"%", sep=""), "End"))+
  theme_pub()

###

fig_seg3.col = 
  dayloopsum %>% filter(variable ==  "v") %>% 

ggplot()+
  geom_pointrange(aes(x = loopprop, y = mean*2, ymin = (mean-ci)*2, ymax = (mean+ci)*2, shape = season2, col = season2), fill = "white", lty = 1, lwd = 0.5, fatten = 5, stroke = 0.4, position = position_dodge(width = 0.1))+
  scale_color_manual(values = c(cola2, colb2))+
 #  scale_fill_manual(values = c(cola2, colb2))+
  scale_shape_manual(values = c(21,16))+
    labs(list(x = "Track stage", y = "Speed (m/hr)", title = "(c)"))+
  scale_x_continuous(breaks = seq(0,1,0.5), labels = c("Start", paste(50,"%", sep=""), "End"))+
  theme_pub()
```

## mindw and stage

```{r}
fig_seg1 = 
dayloopsum %>% filter(variable ==  "mindw") %>% 

ggplot()+
  geom_pointrange(aes(x = loopprop, y = mean, ymin = mean-ci, ymax = mean+ci, shape = season2), fill = "white", lty = 1, lwd = 0.5, fatten = 5, stroke = 0.4, position = position_dodge(width = 0.1))+
#  scale_color_manual(values = c(cola, colb))+
#   scale_fill_manual(values = c(cola2, colb2))+
  scale_shape_manual(values = c(21,16))+
    labs(list(x = "Track stage", y = "Distance to water (m)", title = "(a)"))+
  scale_x_continuous(breaks = seq(0,1,0.5), labels = c("Start", paste(50,"%", sep=""), "End"))+
  #facet_wrap(~variable, scales = "free")+
  theme_pub()

###

fig_seg1.col = 
dayloopsum %>% filter(variable ==  "mindw") %>% 

ggplot()+
  geom_pointrange(aes(x = loopprop, y = mean, ymin = mean-ci, ymax = mean+ci, shape = season2, col = season2), fill = "white", lty = 1, lwd = 0.5, fatten = 5, stroke = 0.4, position = position_dodge(width = 0.1))+
  scale_color_manual(values = c(cola2, colb2))+
#   scale_fill_manual(values = c(cola2, colb2))+
  scale_shape_manual(values = c(21,16))+
    labs(list(x = "Track stage", y = "Distance to water (m)", title = "(a)"))+
  scale_x_continuous(breaks = seq(0,1,0.5), labels = c("Start", paste(50,"%", sep=""), "End"))+
  #facet_wrap(~variable, scales = "free")+
  theme_pub()
```

## export segment figures

### b/w fig
```{r}
pdf(file = "fig_seg_vars.pdf", height = 3, width=7)
library(patchwork)

fig_seg1+fig_seg2+fig_seg3

dev.off()
```

### colour fig

```{r}
pdf(file = "fig_seg_vars.colour.pdf", height = 3, width=7)
library(patchwork)

fig_seg1.col+fig_seg2.col+fig_seg3.col

dev.off()
```


```{r}

#figloop_woods = 
  dayloopdata %>%
    #group_by(season2, loopprop = as.factor(round_any(loopprop, 0.1))) %>%
   # summarise(meanhour = mean(hour)) %>%
  ggplot()+

 geom_smooth(aes(x = loopprop, y = woody.density, col = season2, fill = season2), lty = 1, se = T, lwd = 0.5)
   # geom_violin(aes(x = loopprop, y = woody.density), fill = "grey90", col = "transparent", position = position_dodge(width = 0.95))+
   # geom_boxplot(aes(x = loopprop, y = woody.density, col = season2, fill = season2), position = position_dodge(width = 0.95), lwd = 0.3, notch = F, outlier.size = 0.2, outlier.colour = 1, outlier.shape = 21)+
  scale_color_manual(values = c(cola, colb))+
    scale_fill_manual(values = alpha(c(cola, colb),0.2))+
  scale_x_discrete(breaks = seq(0,1,0.2), labels = c("Start", paste(seq(20,80,20),"%", sep=""), "End"))+
    labs(list(x = "Loop stage", y = "Woody density"))+
  g1
```



# Load speed mod

```{r}
load("ele.mod.speed.RData")
load("ele_with_dem.rdata")
## Fig 3a. Speed ~ temp
#'
ele.speed.temp = ele2 %>% #filter(!id %in% problem_eles) %>%
  mutate(v.pred = predict(mod.speed, newdata = ., scale = "response", allow.new.levels = T), temp = round_any(temp,2)) %>%  group_by(season2, temp) %>%
  summarise(v.mean = mean(v), v.sd = sd(v), n.v = length(v), pred.mean = mean(v.pred, na.rm = T), pred.sd=sd(v.pred, na.rm = T), pred.n = length(v.pred)) %>%
  mutate(v.ci = qnorm(0.975)*v.sd/sqrt(n.v), ci.pred = qnorm(0.975)*pred.sd/sqrt(pred.n))
```

## prepare speed~temp plot all data
### b/w plot
```{r plot_speed_temp}
pdf(file = "fig4speed.pdf", height = 3, width=5)
#'plot begins here

#fig2f = 
  ele.speed.temp %>% filter(temp %in% 15:40) %>% 
  ggplot()+
  geom_smooth(aes(x = temp, y = pred.mean*2, lty = season2), col = 1, lwd = 0.5, fill = "grey")+
 
  geom_pointrange(aes(x = temp, y = v.mean*2, ymin = (v.mean-v.ci)*2, ymax = (v.mean+v.ci)*2, shape = season2), fill = cola3, col = 1, fatten = 5, lty = 1, stroke = 0.6)+
    
 # scale_colour_manual(values = c(cola2,colb2))+
 # scale_fill_manual(values = c(cola1,colb1))+
  scale_shape_manual(values=c(21,24))+
  theme_pub()+
  labs(list(x = "Temperature (째C)", y = "Speed (m/hour)",
            col = "Season", fill = "Season"))

dev.off()
```

### colour plot

```{r plot_speed_temp}
pdf(file = "speed.temp.colour.pdf", height = 4, width=4)
#'plot begins here

#fig2f = 
  ele.speed.temp %>% filter(temp %in% 15:40) %>% 
  ggplot()+
  geom_smooth(aes(x = temp, y = pred.mean*2, lty = season2, fill = season2, col = season2), lwd = 0.5)+
 
  geom_pointrange(aes(x = temp, y = v.mean*2, ymin = (v.mean-v.ci)*2, ymax = (v.mean+v.ci)*2, shape = season2, col = season2), fill = "white", fatten = 5, lty = 1, stroke = 0.6)+
    
    scale_linetype_manual(values = c(2,1))+
 scale_colour_manual(values = c(cola2,colb2))+
 scale_fill_manual(values = c(cola1,colb1))+
  scale_shape_manual(values=c(21,16))+
  theme_pub()+
  labs(list(x = "Temperature (째C)", y = "Speed (m/hour)",
            col = "Season", fill = "Season"))
dev.off()
```
