---
  editor_options:
    chunk_output_type: console
  chunk_output_type: console
---
# Temp and distw fates of incoming eles

# Load libs

```{r load_libs, message=FALSE, warning=FALSE}
#set env
library(plyr);library(dplyr);library(purrr);library(purrrlyr);library(data.table);library(lubridate);library(tidyr)

library(ggplot2)
```

```{r ggplot_theme}
source("libs.R")
source("elecode2.6dataloading.r")
source("ggplot.opts.r")
```

```{r get_data}
#load("ele_after_move.rdata")
```

# Filter data
Filter data to include only:
1. Incoming eles, waterdiff > -500:+500
2. Distwater <= 500
3. All points satisfying both above, and the next 10 points.

```{r temp_fates_data}
#'work on the split ele02 as etf1, ie, ele temp fates 1
ele2$id = ele2$from.idData.rep.1..n.locs.from.....

ele2 = ele2 %>% dlply("id") %>%
  map(function(x) plyr::arrange(x, time) %>% mutate(tempdiff = c(NA, diff(temp, lag = 1)), waterdiff = c(diff(mindw, lag = 1), NA))) %>% bind_rows()

#'split into list of 13
etf1 = ele2 %>% dlply("id") %>% map(function(x) mutate(x, arrival = ifelse(waterdiff < 0 & mindw <=500, "arrival", "other")))

#'split each df into a list by the arrival event
etf2 = etf1 %>% map(function(x) x %>% dlply("arrival"))

#'find the timediff in the arrival events df
for(i in 1:13){
  x = etf2[[i]]$arrival
  etf2[[i]]$arrival$timediff = c(NA, difftime(x$time[-1], x$time[-dim(x)[1]], units = "hours"))
  etf2[[i]]$arrival = etf2[[i]]$arrival %>% filter(timediff > 5)
}

#'bind rows
etf3 = etf2 %>% map(rbind_list) %>% map(function(x) arrange(x, time))
```

```{r id_consecutive_points}
#'find the next 10 points at each arrival event
for(i in 1:13){
  a = which(etf3[[i]]$arrival == "arrival")
  y = unique(sort(a + rep(-10:10, each = length(a))))
  y = y[y>0]
  etf3[[i]] = etf3[[i]][y,]
}

etf3 = etf3 %>% map(function(x) x = x %>% mutate(event = cumsum(arrival == "arrival"), event.inv = event+1))

#etf3 = etf3 %>% map(function(x) x = x %>% dlply("event") %>% map(function(y) y = mutate(y, watertime = difftime(time, time[1]))))
```

```{r water_time_tempchange}
#'bind the lists
etf4 = etf3 %>% map(function(x) x = x %>% group_by(event) %>% mutate(watertime = difftime(time, time[1], units = "hours"), deltadw = mindw - mindw[1], tempchange = temp - temp[1]))

#### Prior to arrival ####
#'make an etf5 for prior to arrival events
etf5 = etf3 %>% map(function(x) x = x %>% group_by(event.inv = ifelse(arrival == "arrival", event, event.inv)) %>% mutate(watertime = difftime(time, time[length(time)], units = "hours"), deltadw = mindw - mindw[length(mindw)], tempchange = temp - temp[length(temp)]))

#'bind rows
etf4 = as.data.frame(bind_rows(etf4))

#'now for pre-arrival
etf5 = as.data.frame(bind_rows(etf5))
```

# Run model

```{r run_glmm_tempchange_watertime,eval=FALSE}
#'run a glmm
library(lme4)
mod.tempchange = lmer(tempchange ~ watertime + distw + season2 + (1|id), data = etf4)

summary(mod.tempchange)
```

# Distw ~ arrival data

```{r distw_watertime}
#'get data

etf.fig =
  etf4 %>%
#  mutate(pred = predict(mod.deltawd, newdata = etf4, allow.new.levels = T, scale = "response")) %>%
  group_by(watertime = round_any(as.numeric(watertime), 0.5), season2) %>%
  summarise(deltawd.mean = mean(deltadw, na.rm = T), deltawd.sd = sd(deltadw, na.rm = T), deltawd.n = length(deltadw), tdmean = mean(tempchange, na.rm = T), tdsd = sd(tempchange, na.rm = T), td.n = length(tempchange)#, predmean = mean(pred, na.rm = T), predsd = sd(pred, na.rm = T), predn = length(pred)
            ) %>%

mutate(deltawd.ci = qnorm(0.975)*deltawd.sd/sqrt(deltawd.n), td.ci = qnorm(0.975)*tdsd/sqrt(td.n)#,#pred.ci = qnorm(0.975)*predsd/sqrt(predn)
       )
```

# ETF inverse: Changes in distance to water prior to arrival

```{r}
etf.fig.inv = etf5 %>%
#  mutate(pred = predict(mod.deltawd, newdata = etf4, allow.new.levels = T, scale = "response")) %>%
  group_by(watertime = round_any(as.numeric(watertime), 0.5), season2) %>%
  summarise(deltawd.mean = mean(deltadw, na.rm = T), deltawd.sd = sd(deltadw, na.rm = T), deltawd.n = length(deltadw), tdmean = mean(tempchange, na.rm = T), tdsd = sd(tempchange, na.rm = T), td.n = length(tempchange)#, predmean = mean(pred, na.rm = T), predsd = sd(pred, na.rm = T), predn = length(pred)
            ) %>%

mutate(deltawd.ci = qnorm(0.975)*deltawd.sd/sqrt(deltawd.n), td.ci = qnorm(0.975)*tdsd/sqrt(td.n)#,#pred.ci = qnorm(0.975)*predsd/sqrt(predn)
       )
```

# Merge data for pre- and post-arrival

```{r}
etf.merged = rbind(etf.fig, etf.fig.inv) %>% filter(watertime >= -12, watertime <= 12)
```

# Distance to water ~ arrival

```{r fig8.distw.arrival}
#'plot fig 6

#cairo_pdf(filename = "fig5distw_arrivalcirc.pdf", height = 4, width = 4, fallback_resolution = 600)

fig5a = ggplot()+
   geom_vline(aes(xintercept = c(0.1,12.5)), lty = 2, col = 1, lwd = 0.2)+
 # geom_smooth(aes(x = watertime, y = predmean, col = season2, fill = season2), lwd = 0.5)+

   geom_hline(yintercept = seq(-500, 2000, 500), lty = 1, col = c("royalblue", rep("grey",5)), lwd = c(1,rep(0.1,5)))+
   geom_bar(data = etf.merged,aes(x = watertime, y = deltawd.mean, fill = season2), stat = "identity", position = "dodge")+

  geom_text(aes(x = 0, y = seq(5e2, 2e3, 5e2), label = c("500m","1000m","1500m","2000m")), vjust = "outward", size = 2)+
  geom_text(aes(x = 0.1, y = -1000, label = "water"), col = "royalblue")+

 # scale_colour_manual(values = c(cola,colb))+
  scale_fill_manual(values = c(cola,colb))+
 # scale_shape_manual(values = c(21, 24))+
  theme_bw()+g1+
  labs(list(x = NULL, y = NULL, title = "A"))+
  xlim(-12,12)+ ylim(-1000, 2e3)+
 coord_polar(start = pi, theta = "x")+
  theme(axis.ticks.y = element_blank(), axis.text.y = element_blank())+
    scale_x_continuous(breaks = seq(-12,12,1))

dev.off()
```

# Temp change ~ arrival

## Get data
```{r}

towtemp$change = c(NA, diff(towtemp$meantempa))

etf4 = merge(etf4, towtemp, by = "hour")

etf.fig.temp =
  etf4 %>% filter(hour %in% 10:16) %>%
#  mutate(pred = predict(mod.deltawd, newdata = etf4, allow.new.levels = T, scale = "response")) %>%
  group_by(watertime = round_any(as.numeric(watertime), 0.5), season2) %>%
  summarise(meanambtempchange = mean(change, na.rm = T), deltawd.mean = mean(deltadw, na.rm = T), deltawd.sd = sd(deltadw, na.rm = T), deltawd.n = length(deltadw), tdmean = mean(tempchange, na.rm = T), tdsd = sd(tempchange, na.rm = T), td.n = length(tempchange)#, predmean = mean(pred, na.rm = T), predsd = sd(pred, na.rm = T), predn = length(pred)
            ) %>%

mutate(deltawd.ci = qnorm(0.975)*deltawd.sd/sqrt(deltawd.n), td.ci = qnorm(0.975)*tdsd/sqrt(td.n)#,#pred.ci = qnorm(0.975)*predsd/sqrt(predn)
       )
```

```{r}
etf5 = merge(etf5, towtemp, by = "hour")

etf.fig.temp.inv =
  etf5 %>% filter(hour %in% 10:16) %>%
#  mutate(pred = predict(mod.deltawd, newdata = etf4, allow.new.levels = T, scale = "response")) %>%
  group_by(watertime = round_any(as.numeric(watertime), 0.5), season2) %>%
  summarise(meanambtempchange = mean(change, na.rm = T), deltawd.mean = mean(deltadw, na.rm = T), deltawd.sd = sd(deltadw, na.rm = T), deltawd.n = length(deltadw), tdmean = mean(tempchange, na.rm = T), tdsd = sd(tempchange, na.rm = T), td.n = length(tempchange)#, predmean = mean(pred, na.rm = T), predsd = sd(pred, na.rm = T), predn = length(pred)
            ) %>%

mutate(deltawd.ci = qnorm(0.975)*deltawd.sd/sqrt(deltawd.n), td.ci = qnorm(0.975)*tdsd/sqrt(td.n)#,#pred.ci = qnorm(0.975)*predsd/sqrt(predn)
       )
```

```{r}
etf.temp = rbind(etf.fig.temp, etf.fig.temp.inv) %>% filter(watertime >= -5, watertime <= 5)
```


```{r}
fig5b =
etf.temp %>%
  ggplot()+
  geom_hline(yintercept = 0, lty = 1, col = "grey", lwd = 0.2)+
  geom_vline(xintercept = 0, lty = 1, col = "grey", lwd = 0.2)+

 # geom_smooth(aes(x = watertime, y = meanambtempchange), col = 1)+
  geom_pointrange(aes(x = watertime, y = tdmean, ymin = tdmean-td.ci, ymax=tdmean+td.ci, col = season2, shape = season2), fatten = 3, stroke = 0.8, fill = "white",position = position_dodge(width = 0.2))+
  scale_shape_manual(values = c(21,24))+
  scale_colour_manual(values = c(cola, colb))+
  labs(list(x = "Time (hours)", y = "Δ temperature (°C)", title = "(b)"))+theme_bw()+g1

```

# 3d surface

## get data

```{r}
etf = bind_rows(etf4, etf5) %>% mutate(phase = ifelse(hour %in% 9:17, "day", "night"))

etf = etf %>% filter(watertime >= -5, watertime <= 5, mindw < 500, tempchange %in% -10:10) %>%
  group_by(watertime = round_any(as.numeric(watertime), 0.5), hour) %>%
  summarise(tdmean = mean(tempchange, na.rm = T), tmean = mean(temp, na.rm = T))


library(reshape2)
z.etf = acast(etf, hour~watertime, value.var = "tdmean")
```


```{r}
library(plot3D); library(RColorBrewer)

cairo_pdf(filename = "fig5b_arrival_hour_temp.pdf", height = 4, width = 4)


#fig5craster =
  ggplot(etf)+

  geom_raster(aes(x = hour, y = watertime, z = tdmean, fill = tdmean))+
  scale_fill_gradientn(colours = rev(brewer.pal(9, "RdBu")))+
  theme_bw()+g1+
  labs(list(x = "Hour", y = "Time to/since arrival (hrs)", title = "C", fill = "Δ Temp"))+
  theme(legend.position = "bottom")

image2D(x = unique(etf$hour), y = unique(etf$watertime), z = z.etf, colvar = z.etf, col =  rev(brewer.pal(9,"RdBu")), ticktype = "detailed", main = "B",clab = "Δ Temp", line = 0, resfac = 15, contour = F, xlab = "Hour",ylab = "Time to arrival (hrs)")
dev.off()
```

## 3d html

```{r}
library(plotly);library(htmlwidgets)

fig5b_3d_html = plot_ly(x = unique(etf$hour), y = unique(etf$watertime), z = ~z.etf, type = "surface", colors = rev(brewer.pal(9, "RdBu")),  name = "Temp") %>% layout(title = "Fig 5b. 3D-interactive",
    scene = list(
      xaxis = list(title = "Hour"),
      yaxis = list(title = "Time ~ arrival"),
      zaxis = list(title = "Temp")
    ), legend = list(z="Temp"))

saveWidget(fig5b_3d_html, "fig5b_3d_interactive.html")
```

# Export figs

```{r}

library(gridExtra)

cairo_pdf(filename = "fig5.pdf", height = 6, width = 8, fallback_resolution = 600)

grid.arrange(fig5a, fig5b, fig5craster, layout_matrix = rbind(c(1,2),c(1,3)))

dev.off()
```
