---
output: html_document
editor_options: 
  chunk_output_type: inline
---

# Prepare data

```{r}
#'load libs and ele data
source("libs.R");source("ggplot.opts.r")
source("ele.move.r")

#'load ssf functions
source("ssf-codes.R")
```


## Set step duration

```{r}
step_duration <- 360

ele.thin <- moveStack(lapply(split(ele.move),thin_split, step_duration, unit = "mins"))
```

## Remove observations without movement

```{r}
#'which records have no movement?
which(unlist(distance(ele.thin)) <= 0)
#'some 2

#'remove them
if (any(unlist(distance(ele.thin)) <= 0)) {
  ele.thin = ele.thin[-which(unlist(distance(ele.thin)) <= 0),  ]
  ele.thin = moveStack(lapply(split(ele.thin),
                                   thin_split, step_duration, unit = "mins"))
}
```

## Raster brick

```{r}
# get raster brick
source("env.stack.r")

```


# Create control locations for each case

```{r}
ele.angles <- prepare_angle_dist(ele.thin)

# Inspect the data
summary(ele.angles)
```

## Empirical distances and turning angles

```{r }
#plot observed distances and headings
par(mfrow = c(1, 2))
hist(ele.angles[,"dist"], breaks = 20, main = "",
     xlab = "Distance (m)")
hist(ele.angles[, "rel.angle"], breaks = seq(-pi, pi,
                                             len=11), main="", xlab="Relative angle (radians)")
```


## Fit distributions to distances

```{r}
# get the distributions, plot them over the data
fexp <- fitdistr(ele.angles[, "dist"], "exponential")
#fgam <- fitdistr(ele.angles[, "dist"], "gamma")
#flogn <- fitdistr(ele.angles[, "dist"], "lognormal")

par(mfrow = c(1,1))

#'histogram distances
hist(ele.angles[,"dist"], breaks = 150, prob = TRUE, xlim = c(0, 8000), ylim = c(0, 2e-3), xlab = "Step length (m)", main = "")

#'if the data were fitted to an exponential distribution, their exponent would be fexp$estimate (an exponent of values of x, here 0 -- 5000). this is visualised below
plot(function(x) dexp(x, rate = fexp$estimate), add = TRUE, from = 0, to = 5000, col = 2)
```

## Fit distribution to turning angles

```{r}
#'some interesting distribution fitting...why?
fkappa <- est.kappa(ele.angles[,"rel.angle"])
fkappa

#'plot to see
hist(ele.angles[, "rel.angle"], prob = TRUE, main = "", xlab = "Turning angle")
plot(function(x) dvonmises(x, circular(0), kappa=fkappa), add = TRUE, from = -pi, to = pi, col = "red")
```

## Change to a flat CRS

Currently, the code for generating alternative steps assumes a suitable "flat" projection, so we pick the appropriate UTM projection for the study area.

```{r}
#'create the crs we need, this is zone 36S
utm_crs <- CRS("+proj=utm +zone=36 +south +datum=WGS84 +units=m +no_defs +ellps=WGS84 +towgs84=0,0,0")
```


## Sampling for alternatives

```{r}
#sample 5 alternative points at each point, with some methods, here the exponential, and a theta of half the exponent estimate

my_K <- 5

set.seed(3)
ele.ssf <- prepare_ssf_steps(ele.thin,
                                  method = "exponential", K = my_K,
                                  theta = fexp$estimate / 2,
                                  crs = utm_crs)
```



# Extract the raster information at step locations

```{r}
envmat <- extract(env, spTransform(ele.ssf, env@crs))

```

# Add the information to the existing data set

```{r}
#Caution - do the cbind on the @data part only
ele.ssf@data <- cbind(ele.ssf@data, envmat)
head(ele.ssf)
# take a look at the data
```



```{r add_time}
#' add hour: allows us to model variation in step lengths, turning angles, and preference for environmental conditions as a function of time of day

ele.ssf$hour <- hour(ele.ssf$date) + minute(ele.ssf$date) / 60

#'plot to check preferences
par(mfrow = c(1, 2))
boxplot(dist ~ floor(hour), data=subset(ele.ssf,used == 1), xlab = "Time of day", ylab = "Distance (m)")

boxplot(cos(rel.angle) ~ floor(hour), data = subset(ele.ssf, used == 1), xlab = "Time of day", ylab = "cos(turning angle)")

par(mfrow = c(1, 1))
```


```{r}
# Remove NA's
ele.ssf <- ele.ssf[complete.cases(ele.ssf@data),]
```


# Collinearity

Check to see if some variables are correlated, something which one would expect. 

```{r check_collinearity}
#round(cor(ele.ssf@data[, c("water","trees","temp")]), 2)

#'no variables are strongly correlated

```


## Picking a variable: conditional logit regression

```{r conditional logit regression, eval=FALSE, include=FALSE}
#m1_exp_water <- clogit(used ~ cos(rel.angle) + dist + water_dist + strata(stratum), data = ssf_cilla)
#m1_exp_elev <- clogit(used ~ cos(rel.angle) + dist + elev + strata(stratum), data = ssf_cilla)
#AIC(m1_exp_water)
#AIC(m1_exp_elev)
 
# So we pick elev, because it by itself explains the movement better has a lower AIC. 
```

# Calculate step selection function

```{r ssf}
#clogit is the conditional logistic regression from the package survival the response is used, coded as 1, not used is 0. strata is a novel term.

#'model 1. use the term stratum (here strata), as a sort of random effect
library(mclogit)
m_1 <- clogit(used ~ cos(rel.angle) + sqrt(waterraster):pbs(hour, df = 4, Boundary.knots = c(0,24)) + trees_interpolate:pbs(hour, df = 4, Boundary.knots = c(0,24)) + strata(stratum), data = ele.ssf)

a = predict(m_1, newdata = ele.ssf)

ele.ssf$pred = a

ele.ssf2 = data.frame(ele.ssf)

ele.ssf2 = ele.ssf2 %>% group_by(hour)
summary(m_1)

by1k = function(x) x/1e3

png(filename = "ssf_fig.png", res = 300, height = 2000, width = 1600)
visreg(m_1, xvar = "waterraster", by = "hour", breaks = c(5, 10, 15, 20), xtrans = by1k, xlab = "Distance to water (km)", ylab = "Selection strength", rug = T, strip.names = c("10 - 15", "15 - 20", "20 - 05", "05 - 10"), line.par = list(col = 1, lwd = 0.5), partial = F)
dev.off()
```

```{r}
png(filename = "ssf_fig_trees.png", res = 300, height = 2000, width = 3200)
visreg(m_1, xvar = "trees_interpolate", by = "hour", breaks = c(5, 10, 15, 20), strip.names = c("10 - 15", "15 - 20", "20 - 05", "05 - 10"), xlab = "Woody density", ylab = "Selection strength", rug = T, line.par = list(col = 1, lwd = 0.5), partial = F)
dev.off()
```


# Model selection
Model selection is a vast topic. Here just use stepwise backward selection based on AIC

```{r eval=FALSE, include=FALSE}
#'check aic for updated models
m_2 <- update(m_1, .~. - slope)
AIC(m_1)
AIC(m_2)
summary(m_2)
```

# Model checking: serial autocorrelation

see *Forester et al. 2009 Ecology 90:3554–3565*

Calculate the deviance residuals for each stratum (i.e., the sum of the residuals for the case and all associated controls).

```{r}
#'checking the residuals
ele.ssf$residuals <- residuals(m_1, type = "deviance")

resid_df <- group_by(ele.ssf@data, date)
resid_df <- summarise(resid_df, residuals = sum(residuals))

#
resid_df$group <- 1

#' Fit an intercept-only mixed-effects model using lme() from the nlme package.
library(nlme)
rm1 <- lme(residuals ~ 1, random = ~ 1 | group,
           data = resid_df)

plot(ACF(rm1), alpha = 0.05)
```

There seems to be some residual temporal autocorrelation. One effect of residual temporal autocorrelation is extreme p-values, but it may also cause bias in parameter estimates. In the case of autocorrelation, increasing step-length is a good idea, and the p-values aren't the most trustworthy.


# Model evaluation

 - R2 is low. Always is.
 - Not yet clear what a good performance index would be.
 - Consider this an exercise.
 
Here: we contrast for used and alternative steps the "Proportion of values less than or equal to the current value".

 - Cross-validation
 - split steps in e.g. 5 folds (long stretches better)
 - leave each fold out, refit and predict to left-out fold


```{r model eval}
#'get model predictions
ele.ssf$pred <- predict(m_1)
by_stratum <- group_by(ele.ssf@data, stratum)
#'assign a prediction rank using a windowed rank function
by_stratum <- mutate(by_stratum, pred_rank = cume_dist(pred))

#'plot a histogram of...?
hist(subset(by_stratum, used == 1)$pred_rank, xlab = "AUC", main = "", prob = TRUE)
#'Mean AUC across steps
mean(subset(by_stratum, used == 1)$pred_rank)
```

# Interpretation

 - Map preference
 - Variable importance (see SDM this afternoon)
 - Response functions (only interesting for time varying parameters)


## Map habitat preference

The raster prediction function assumes that all environmental layers are respresented in one raster stack.

```{r get_predictions}
# required for prediction by clogit model, but does not affect result
stratum <- extract_strata(m_1)[1]

pred_df_map <- data.frame(getValues(env))
pred_df_map$stratum <- stratum
pred_df_map$dist <- 0
pred_df_map$rel.angle <- 0
```

```{r map_predictions}
m1_pred <- predict(m_1, newdata = pred_df_map)
m1_pred_raster <- raster(env, 1) # Create raster layer
m1_pred_raster[] <- m1_pred
plot(m1_pred_raster)
```

## Overlay eles

```{r }
am105 <- split(ele.move); am105 = am105$AM105
plot(m1_pred_raster)
points(am105, cex = 0.1)
```


# Time-varying preference for water

```{r time and water}
boxplot(waterraster ~ floor(hour), data = subset(ele.ssf, used == 1), xlab = "Time of day", ylab = "Distance to water (m)", notch = T, ylim = c(0,4e3))

#'some sort of gam

#'use a bam
library(mgcv)
m_time <- bam(used ~ sqrt(waterraster) + trees_interpolate+ waterraster:hour, data = ele.ssf, family = binomial)


summary(m_time)
```

#Predictions with the model: response function

```{r}
pred_data <- data.frame("stratum" = stratum, dist = 0,
                        trees_interpolate =  , waterraster =  2000, rel.angle = 0, hour =
                          seq(0, 24, len = 101))
pred_time <- predict(m_1, newdata = pred_data,
                     se.fit = TRUE, scale = "response")
upper <- pred_time$fit + 1.96 * pred_time$se.fit
lower <- pred_time$fit - 1.96 * pred_time$se.fit
```

# Predictions with the model: response function

```{r}
par(mfrow = c(1, 1))
plot(pred_data$hour, pred_time$fit, type = "l",
     ylim = range(c(upper, lower)),
     ylab = "Preference")
lines(pred_data$hour, upper, lty = 2)
lines(pred_data$hour, lower, lty = 2)
abline(h = 0, lty = 3)
```

#take a look at the preference against time of day

#### Simulating Cilla ####
 Simulating with the model
 =========================
 We start at the first recorded position of the individual "Cilla"
coordinates(spTransform(subset(ssf_data_exp, id ==
                                 "Cilla" & used == 1), utm_crs))[1, ]
start_xy <- c("x" = 384885.8, "y" = 7240399.0)
 We assume that the last step was a move eastwards
start_angle <- 0


 Simulating with the model
 =========================
 We arbitrarily pick the first stratum in the model - which one we pick does not affect the
 results, but we need to supply a stratum for the predict.coxph function

stratum <- extract_strata(m_3)[1]
set.seed(2)
sim_1 <- simulate_ssf(m_3, map = env, nsteps = 50,
                      K = 200, method = "exponential", theta = fexp$estimate / 2,
                      start_xy = start_xy, start_angle = start_angle,
                      stratum = stratum, crs = utm_crs)

 Plotting the simulated steps
 =====================================================
sim_1 <- spTransform(sim_1, env@crs)
plot(crop(raster(env, "elev"), extent(sim_1) + 5000))
lines(coordinates(sim_1), col ="red")

 Comparing with the real first 50 steps of Cilla
 =====================================================
cilla_steps <- subset(ssf_data_exp, id == "Cilla" &
                        used == 1)
cilla_steps <- spTransform(cilla_steps, utm_crs)
elev <- raster(env, "elev")
elev_crop <- crop(elev, extent(cilla_steps[1:50, ]) +
                    15000)

 Comparing with the real first 50 steps of Cilla
 =====================================================
plot(elev_crop)
lines(coordinates(cilla_steps)[1:50, ], col = "red")
lines(coordinates(sim_1))

#### How the sim does ####
 not very well, Cilla has gone over a mountain, which seems unlikely.
 now, the two points on either side of the mountain crossing are very reasonable
 but the intervening section is not
 this is because the function is looking for the next step, which needs to be of
 a certain elevation, and the opposite side of the hill satisfies that
 the path then needs to be taken into account. this could be done if the ascent,
 or if the energy required in completing each segment of the path were to be added
 to the sim.
 a least cost path will most probably have only a single solution, but it does
 weigh the cost of moving toward a high avoidance region and moving to avoid it
 yet reach a point beyond.
 A utility based point selection function is another option. See on Monday.
 What's circuitscape?

 Test method with simulated data
 ====================================
 Question: Is the method able to recover the "true" parameter values
 from data?

 Approach:

 - Simulate data with a known model
 - Fit method to simulated data
 - Compare fitted parameters to those of the known model


 Prepare parameters and environmental data
 ==========================================
 We pick the model m_3 to represent Truth
generating_model <- m_3
 Setting of parameter values

start_xy <- c("x" = 384885.8, "y" = 7240399.0)
start_angle <- 0
stratum <- extract_strata(m_3)[1]

 Simulate data with a known model
 =====================================================
set.seed(4)
if(file.exists(paste0("simulated_path_", step_duration, "_", my_K, ".rds"))) {
  simulated_path <- readRDS(paste0("simulated_path_", step_duration, "_", my_K, ".rds"))
} else {
  simulated_path <- simulate_ssf(generating_model,
                                 map = env, nsteps = 500, K = 1000, method =
                                   "exponential", start_xy = start_xy, start_angle =
                                   start_angle, stratum = stratum, theta = fexp$estimate / 2,
                                 crs = utm_crs, verbose = FALSE)
  saveRDS(simulated_path, file = paste0("simulated_path_", step_duration, "_", my_K, ".rds"))
}

 Make a move object from the simulated data
 ===========================================
start_date <- min(timestamps(buffalo))
simulated_data <- move(
  x = coordinates(simulated_path)[, 1],
  y = coordinates(simulated_path)[, 2],
  time = seq(start_date, len = NROW(simulated_path),
             by = paste(step_duration, "mins")),
  proj = simulated_path@proj4string)

 Plot simulated data
 =====================================================
simulated_data_utm <- spTransform(simulated_data,
                                  env@crs)
elev_crop <- crop(elev, extent(simulated_data_utm) +
                    5000)

 Plot simulated data
 =====================================================
plot(elev_crop)
lines(coordinates(simulated_data_utm))

okay, so this simulated line looks just as weird, the animal is crossing
rivers and boundaries that it normally doesn't. take a model where you know the params
run a sim, get the parameters again, and they should be the same.

 Analyse simulated data: thinning and splitting
 =====================================================
thinned_data_sim <- thin_split(simulated_data, step_duration,
                               units = "mins")
angle_dist_sim <- prepare_angle_dist(thinned_data_sim)

hist(angle_dist_sim[, "dist"], prob = TRUE)

fexp_sim <- fitdistr(angle_dist_sim[, "dist"], "exponential")
fkappa_sim <- est.kappa(angle_dist_sim[, "rel.angle"])

 Analyse simulated data: prepare SSF steps and add environmental information
 =====================================================
set.seed(2)
ssf_test_data <- prepare_ssf_steps(thinned_data_sim,
                                   method = "exponential", K = my_K, theta = fexp_sim$estimate / 2,
                                   crs = utm_crs)
envmat <- extract(env, ssf_test_data)
ssf_test_data@data <- cbind(ssf_test_data@data,
                            envmat)

 Analyse simulated data: fit the model
 =====================================================
m_test_exp <- clogit(used ~ cos(rel.angle) + dist + elev +
                       mean_NDVI + strata(stratum),
                     data = ssf_test_data)
coef(generating_model)
coef(m_test_exp)

#okay, the two seem to be very similar.

summary(m_test_exp)
 For the movement-related parameters, we need to take the parameters of
 the proposal steps into account
 the sim helps establish how many control points one needs

 True rate
true_rate <- fexp$estimate / 2 - coef(generating_model)["dist"]
 Estimated rate
est_rate <- fexp_sim$estimate / 2 - coef(m_test_exp)["dist"]

 True mean distance
1/true_rate
 Estimated mean distance
1/est_rate

knowledge of an area could be a predictor if HR tendency was a factor.

 From preference to utilisation maps
 ==================================
 There is a fast method if we have a symmetric jump kernel, e.g. exponential, and no effect of step angles.
 Barnett, A. & Moorcroft, P. (2008) Analytic steady-state space use patterns and rapid computations in mechanistic home range analysis. Journal of Mathematical Biology, 57, 139–159.

